Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-07-03 paper/ms.tex
update ms

A few competing explanations for what causes the complex quasiperiodic
variability are shown in Figure~\ref{fig:f1}, along with some
representative light curves.  All young M-dwarfs are spotted, which
produces flux variations over characteristic timescales of the
rotation period, $P_{\rm rot}$, and its half-harmonic, $0.5\,P_{\rm
rot}$.  The observed dips occur over durations as short as
$0.05\,P_{\rm rot}$; a ``starspot-only'' scenario can be discarded for
any object with sufficiently sharp dips
\citep{2017AJ....153..152S,2021MNRAS.500.1366K}.   The timescales and
amplitudes of the flux variations instead require sharp geometries
with material extrinsic to the stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  The
``clump'' scenario invokes opaque dust clumps orbiting near the Keplerian
corotation radius, which periodically eclipse the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  The
``prominence'' scenario invokes long-lived condensations of cool and dense
marginally-ionized gas, embedded within the
hotter corona, that would be centrifugally supported near corotation
\citep{1989MNRAS.238..657C,2019MNRAS.482.2853J,2022MNRAS.514.5465W}.
Such structures are analogous to quiescent prominences and filaments
observed on the Sun, though at much larger relative distances from the
stellar surface.
A final posibility that has been suggested is one of a ``screen'', in
which the inner wall of a quiescent circumstellar disk blocks a
portion of the stellar surface to produce sudden dips whenever spots
come into view \citep{2019ApJ...876..127Z}.

In our view, the dust clump and prominence hypotheses seem most
plausible.  They are qualitatively similar, except one
invokes opacity from dust, while the other invokes opacity from gas,
likely bound-free transitions in hydrogen or perhaps a
molecular opacity.  The screen scenario seems inconsistent with the
degree of observed periodicity, the lack of infrared excess, and the
observed lifetime of CQVs extending an order of magnitude longer than
the $\approx$$10^7$ year timescale typically quoted for the primordial
disk to disperse.  However, unambiguous evidence 
has yet to be acquired in support of any scenario.  Such evidence
might include a spectroscopic detection of silicate 10\,$\mu$m dust
absorption during a dip, or perhaps detection of transient
Balmer-line absorption during the dips, similar to observations made in
systems such as AB~Dor \citep[see the review
by][]{1999ASPC..158..146C}.

@@

We visually assessed whether the objects found using the Fourier
(Section~\ref{subsec:fourier}) and dip-counting
(Section~\ref{subsec:counting}) techniques were consistent with
expectations for CQVs by assembling the data shown in
Appendix~\ref{app:vetting}.  We labelled a star as a ``good'' CQV if
at least one TESS sector showed what we viewed as the unambiguous
signatures of the class (short period; at least three dips or else
otherwise oddly-shaped dips; relative stability over a timescale of
30\,days).  We also noted stars that we thought could be CQVs, but
that were more ambiguous with a ``maybe'' flag.

Broadly speaking, the most common false positives for both the Fourier
and dip-counting techniques were eclipsing binaries, spot-induced
variability from rapid rotators, and variability from neighboring,
off-target stars.  Typical false postive rates from our dip-counting
pipeline were 5:1, with \nuniqdipflagged\ unique stars flagged, and
about 20\% being labelled either ``good'' or ``possible'' CQVs; for
the Fourier pipeline, {\bf the rate was X:Y, with ZZZZ unique stars
flagged, and NN\% being labelled ``good'' or ``possible'' CQVs}.

@@

We determined the stellar effective temperature and radii through SED
fitting; we then estimated the masses by interpolating against
the PARSEC v1.2S models \citep{2012MNRAS.427..127B,2014MNRAS.444.2525C}.
%the ``magnetic'' models of \citet{2016A&A...593A..99F}.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f2.pdf}
    \vspace{-0.7cm}
		\caption{
      {\bf CQVs found in the TESS 2-minute data.}
      Phased TESS light curves over one month are shown for \ngoods\
      CQVs.  Gray are raw 2-minute data; black bins to 300 points per
      cycle.  Each panel is labelled by the TIC identifier, the TESS
      sector number, and the period in hours.  Objects are ordered
      such that sources with the most TESS data available are on top
      (see Section~\ref{sec:catalog}).
		}
		\label{fig:cqvs}
	\end{center}
\end{figure*}

@@

As is typical for CQVs \citep{2017AJ....153..152S}, most CQVs in our
catalog did not show infrared excesseses in the WISE1-4 bands.
Visually inspecting the SEDs of our entire \ncpvsfound\ star sample,
we labelled two objects as having ``good'' infrared excesses (both W3
and W4), and three as ``possible'' infrared excess.

The two ``good'' IR excesses belonged to ``candidate'' CQVs
TIC~193136669 (TWA~34) and TIC~57830249 (TWA~33).  Both are in the TW
Hydrae association ($\approx$10\,Myr), and have relatively long
periods of 38\,hr and 44\,hr respectively.  In our initial labelling,
we labelled both as ``maybe'' CQVs because the dips in their light
curves did not show the rigid periodicity typical of CQVs;  their
periods were also relatively long.  After labelling them, inspection
of further sectors clarified that both sources are dippers (see plots
in Appendix~\ref{app:vetting}).  In addition, TIC 193136669 has a cold
molecular disk based on observed 1.3\,mm continuum emission and
resolved Keplerian $^{12}{\rm CO}(2-1)$ emission
\citep{2015A&A...582L...5R}.  It was also labelled a dipper by
\citet{2022ApJS..263...14C}; we agree with their designation, but
nonetheless leave it in our catalog as an indication of possible
ambiguities in our search process.  TIC 57830249 (TWA~33) also has
previously detected 1.3\,mm continuum emission
\citep{2015A&A...582L...5R}, suggestive of cold dust grains being
present.  It therefore also a dipper that snuck its way into our
``maybe'' pile.
Possible evolutionary connections between CQVs and dippers are
highlighted in Section~\ref{subsec:discdippers}.

Our three ``possible'' infrared excesses were TIC~405910546,
289840926, and 244161191.  After a literature search, we concluded
that none have clear evidence for the presence of a disk.
TIC~405910546, in LCC, shows a unique TESS light curve, reminescent of
a $P$=38\,hr singly-eclipsing binary, except with additional
substructure during each eclipse that resembles the CQV phenomenon
more than any other variability of which we are aware.  TIC~289840926
($\beta$~Pic moving group, $P$=4.8\,hr), show what we believe is a
clear CQV signals, but has no definitive evidence for a large, dusty
disk.  TIC~244161191 (hilariously, TOI-278), in Columba, also has no
definitive evidence for a large disk.  It is however
``multi-periodic''---in addition to the 7.17\,hr CPV signal, this
source shows a superposed 8.39\,hr sinusoidal signal, probably from an
unresolved neighboring star.

@@

Figure~\ref{fig:evoln} shows ``before'' and ``after'' views of 27 CQVs
for which TESS 120-second cadence observations were available at least
two years apart.  Such a baseline was available for 32 of the
confirmed \ngoods\ CQVs in our catalog; for plotting purposes we show
the brightest 27.
Of these 27 CQVs, a few show clear signs of the phenomenon in one
sector, and marginal or non-existence signs in the other.
While there is some subjectivity in this assessment, to our eyes cases
for which at least one sector would be flagged as ``ambiguous''
include
TIC~368129164 (Sector 23 might be labelled an EB),
TIC~177309964 (Sector 38 would be simply a rotating star),
TIC~404144841 (Sector 38 looks like a rotating star),
TIC~201898222 (Sector 3 looks like a rotating star),
TIC~144486786 (Sector 32 might be an RS CVn),
and
TIC~38820496 (Sector 28 might be an RS CVn).
TIC~193831684, assessed on a single-sector basis, would probably be
labelled an eclipsing binary---in fact, \citet{2021ApJ...912..123J}
implicitly have already given this source such a label.  However, based on the
the shape evolution between Sectors 13 and 39, it is a CQV.
One could easily also imagine in cases like TIC~206544316 that if
observed at lower signal to noise, the drastic shape evolution would
not be appreciated.  We emphasize that the periods themselves were all
stable to $<$0.1\%.  Broadly speaking, this shape evolution suggests
that CQVs have an ``on/off'' duty cycle of $\approx$75\%.  This type
of correction is likely worth including in population-level estimates
of how intrinsically common CQVs are in the low-mass stellar
population \citep[e.g.][]{2022AJ....163..144G}.

@@

The average period, determined by measuring the PDM peak period over
each sector independently, was $\langle P \rangle = 18.5560$\,hr.  The
range between the maximum and minimum sector-specifics periods was
measured to be about one minute.  Based on this range, the star's
period is stable to at least one minute ($\pm$0.017\,hr) over the
three year baseline.  However, in detail, a period shift of
$\pm$1\,minute would yield major phase drifts over the baseline; that
time interval corresponds to roughly 1/1000$^{\rm th}$ of a period,
and we have observed 1500 cycles.  The achievable period precision,
$\sigma_{\rm P}$, can be estimated as
\begin{equation}
  \sigma_{\rm P} = \frac { \sigma_{\rm \phi} P } { N_{\rm baseline} },
\end{equation}
for $N_{\rm baseline}$ the number of cycles in the observed baseline
and $\sigma_{\phi}$ the phase precision with which any one feature
(e.g.~a dip, or the overall shape of the sinusoidal envelope) can be
tracked.  Assuming $\sigma_{\rm phi} \approx 0.01$ yields an expected
period precision $\sigma_{\rm P} \approx 0.45 {\rm sec} \approx
1.2\times 10^{-4}\,{\rm hr}$.  By visually fine-tuning over a grid in
period, we eventually found $P=18.5611 \pm 0.0001$\,hr, which seems to
track certain features in the LP 12-502 light curve over its entire
baseline.

@@

The second dip complex during cycles 248-315 shows the most
substructure.  During e.g.~cycles 283-298, this single complex shows
six local minima.  The deepest dip is very sharp: it shows a flux
exercusion of 3.5\% over about 22 minutes (0.02\,$P$), which is the
steepest slope exhibited anywhere in this object's remarkable dataset.
After the sharp dip, there is a roughly exponential fall-off spanning
about a quarter of a period, punctuated by coherent local minima and
maxima which in detail (Appendix~\ref{app:lp}) have slightly
longer periods than the sharp dip.  The sharp leading dip only
decreases during a sudden state-switch at BTJD 2030.7 (cycles
299-315), which happens immediately during a flare
(Figure~\ref{fig:lplc}).  The trailing dips remain afterwards.

Sectors 53--58 (cycles 1233-1481) are comparitively tame; they showed
only four to six dips per cycle.  Some dips remain stable in depth and
duration over this five month interval.  Other dips grow, like the one
at $\phi = +0.06$ between cycles 1499 and 1481.  Other dips, such as
the one at $\phi = +0.12$ in cycles 1233-1264, disappear entirely.
The most dramatic state switch occurs during cycles 1233-1247, when a
large dip ``switches'' from a depth of 3\% and duration of 3 hours to
a depth of 0.3\%, and a duration of 1 hour.

@@

% 16 px vs 310 px for a full phase
%Based on our SED fitting $R_\star=0.369\,R_\odot$
%star, and from the isochrone mass, $M_\star=0.215\,M_\odot$ 
{\sc Dip durations}---The shortest timescale for any of the {\it
individual} LP 12-502 dips seems to be $\approx$0.06\,$P$ $\approx$
1.08\,hr.  In comparison, using the stellar radius and mass derived in
Section~\ref{subsec:starprops}, the characteristic timescale $T_{\rm
dur} \equiv R_\star P / (\pi a)$ for the transit of a point-source at
corotation is 1.03\,hr ({\bf uncertainties?}).  This means that while
some of the LP 12-502 dips are sufficiently long to require structures
that are extended in orbital azimuth, the durations of other dips are
consistent with effective radii for the occulting material $R_{\rm
eff} \ll R_\star$ .
We also emphasize that this implies $a/R_\star \approx 5.8$ for this
material, and so the analogous timescale at the stellar surface is
about six times slower.

@@

\subsection{Dust, or gas?}

We believe that the most likely scenarios to explain the overall CQV phenomenology
are the dust clump scenario, or the
prominence scenario (Figure~\ref{fig:f1}).
Both invoke clumpy material that would be centrifugally supported at the corotation radius.
The prominence idea has a deeper history,
based first on analogy with quiescent prominences/filaments
observed to remain stable in the solar corona for up to a few weeks.
In an extrasolar context,
spectroscopic detections of transient Balmer- and
resonance-line absorption seen for stars such as AB~Dor and
Speedy~Mic \citep[e.g.][]{1989MNRAS.238..657C,1993MNRAS.262..369J,2006MNRAS.365..530D,2016MNRAS.463..965L}
led to the interpretation that the data could best be explained by
cold, minimally ionized hydrogen clouds scattering chromospheric emission from the star,
producing the observed spectroscopic line variations \citep[see][]{1989MNRAS.238..657C}. 
The mechanical stability of such gas configurations has been
demonstrated \citep{2000MNRAS.316..647F,2022MNRAS.514.5465W}, and generally, the interpretation
of this class of observations seems reasonably secure.

@@

%TODO: improve this to make it less jargony
The difference between the prominence and dust clump scenarios
is essentially only in whether the occulting material of
interest is neutral hydrogen, or dust.  In the phrasing of the
``frozen flux'' condition of ideal rigid field magnetohydrodynamics,
the tendency of both to become trapped at the corotation radius in the
equatorial plane is tied to how of the four relevant forces (gravity,
Lorentz, inertial Coriolis, and inertial centrifugal), the Lorentz and
Coriolis only act perpendicular to field lines, while gravity and the
centrifugal force are in balance at $R_{\rm cr}$
\citep[see][Sec.~2]{2005MNRAS.357..251T}.
The magnetic field strength is only relevant in this formulation of
the system in that we must have $R_{\rm sonic} < R_{\rm cr} < R_{\rm Alfven}$ in order for 
closed loops to exist that can support prominences 
\citep{2019MNRAS.482.2853J} {\bf lgb todo: verify!}.
In detail however, whether such magnetic fixed points lie in the equatorial
plane, or elsewhere, depends on the star's magnetic field geometry
\citep{2023MNRAS.518.4734S}.

@@

\subsubsection{Does the CQV lifetime inform the mechanism?}
Independent of the microphysical basis for gas absorption, the
prominence scenario may present another issue: why don't stars in the
Hyades and Praesepe show the phenomenon?
In a broad sense, the prominence scenario argues that an intrinsic
stellar process is key for generating the CQV phenomenon.
If this were true, then stars of identical mass, size, and rotation
periods might be expected to show the phenomenon in equal number,
simply because the stellar dynamo is generated by fluid
motions inside the star, which should be identical for stars of the
same mass, size, and rotation period.

Between ages of $\approx$100\,Myr (AB Dor, Pleiades, Psc-Eri) and
$\approx$700\,Myr (Praesepe and Hyades), a 0.3\,M$_\odot$ star will
shrink by $\approx$10\%, from $\approx$0.33\,R$_\odot$ to 
$\approx$0.29\,R$_\odot$ \citep[e.g.][]{2014AJ....147...82C}.
This is not a particularly drastic size change.
Similarly, at 700\,Myr many M-dwarfs have comparably rapid rotation periods 
to those at 100\,Myr \citep[e.g. Figure~13 of][]{2022AJ....164...80R};
the mean population does show evidence for spin-down after the
PMS contraction finishes, but there are still a large number of 
700\,Myr M-dwarfs with $P_{\rm rot}\lesssim$2\,days, beyond
which the CQV phenomenon becomes less common.

Broadly speaking -- the absence of old CQVs seems to be
more plausibly linked to a gradual depletion of dusty detritus from the
planet formation era than it does to the changing stellar properties
as the stars gradually finish their pre-main-sequence evolution, and then
continue to spin rapidly for many hundreds of millions of years.

@@

It is very tempting to try to interpret features of the
CQV light curves in terms of the possible presence of close-in
exoplanets, or even planetesimals.  Rocky planets this
young would likely have molten global magma oceans analogous to those that
existed on the Earth and Moon \citep[see][]{2022arXiv220310023L}, and thus
would being undergoing significant outgassing and atmospheric escape.  A
scenario in which close-in rocky planets serve as a possible source of
dust for the clouds is not entirely implausible, {\it a priori}.
%Frankly though, it remains a subject of speculation.

@@

It is amusing that for the two candidate CQV sources identified in
Section~\ref{subsec:irexcess} as having outlying SEDs (TICs~193136669
and TIC~57830249), the identification of the infrared excess is part
of what prompted our eventual labelling of these sources as dippers.
It is also worth emphasizing that our labelling of e.g.~TIC~57830249
was based on a single sector of TESS data (Sector~36), when its
behavior is relatively periodic, with dip depths of a few percent.
However in other TESS sectors (e.g. Sector~10), this source looks
completely different, varying in apparent flux by a factor of two,
which hardly any discernable periodicity at all.

@@

It is challenging to interpret the significance of such apparent
regularities;  the objects exhibit such a wide range of variability
that for seemingly any ``regularity'' one might posit, it is easy to
come up with counter-example objects which do not follow the trend.
In other words, these half-cycle CQVs could simply be a
product of the human tendency to pattern match; alternatively, they
might yield some important physical signifiance.

@@

Finally, Figure~\ref{fig:lpriver0} shows ``river plots'' of the
same data, split into very similar intervals:
the Sector 18-19 data, 25-26 data, 53 data, and 58-59 data.
State changes are evident in these plots whenever there is a sudden
change in color.
{\bf todo: fix missing data to be different from the flare color}

>>>>

A few competing explanations for what causes the complex quasiperiodic
variability are shown in Figure~\ref{fig:f1}, along with some
representative light curves.  All young M-dwarfs are spotted, which
produces flux variations over characteristic timescales of the
rotation period, $P_{\rm rot}$, and its half-harmonic, $0.5\,P_{\rm
rot}$.  The observed dips occur over durations as short as
$0.05\,P_{\rm rot}$; a ``starspot-only'' scenario can be discarded for
any object with sufficiently sharp dips
\citep{2017AJ....153..152S,2021MNRAS.500.1366K}.   The timescales and
amplitudes of the flux variations instead require sharp geometries
with material extrinsic to the stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  The
``clump'' scenario invokes opaque dust clumps orbiting near the
Keplerian corotation radius, which periodically eclipse the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  The ``prominence''
scenario invokes long-lived condensations of cool and dense
marginally-ionized gas, embedded within the hotter corona, that would
be centrifugally supported near corotation
\citep{1989MNRAS.238..657C,2019MNRAS.482.2853J,2022MNRAS.514.5465W}.
Such structures are analogous to quiescent prominences and filaments
observed in the solar corona \citep[see e.g.][]{2015ASSL..415.....V},
though at much larger relative distances from the stellar surface.  A
final possibility that has been suggested is one of a ``screen'', in
which the inner wall of a quiescent circumstellar disk blocks a
portion of the stellar surface to produce sudden dips whenever spots
come into view \citep{2019ApJ...876..127Z}.

The dust clump and prominence hypotheses seem most plausible.  They
are qualitatively similar, except that one invokes opacity from dust,
while the other invokes opacity from gas, likely bound-free
transitions in hydrogen or perhaps a molecular opacity.  The screen
scenario seems inconsistent with the degree of observed periodicity,
the lack of infrared excess, and the observed lifetime of CQVs
extending an order of magnitude longer than the $\approx$$10^7$ year
timescale typically quoted for the primordial disk to disperse.
However, unambiguous evidence has yet to be acquired in support of any
scenario.  Such evidence might include a spectroscopic detection of
silicate 10\,$\mu$m dust absorption during a dip, or perhaps detection
of transient Balmer-line absorption during the dips, similar to
observations made in systems such as AB~Dor \citep[see the review
by][]{1999ASPC..158..146C}.

@@

We visually assessed whether the objects found using the Fourier
(Section~\ref{subsec:fourier}) and dip-counting
(Section~\ref{subsec:counting}) techniques were consistent with
expectations for CQVs by assembling the data shown in
Appendix~\ref{app:vetting}.  We labeled a star as a ``good'' CQV if
at least one TESS sector showed what we viewed as the unambiguous
signatures of the class (short period; at least three dips or else
otherwise oddly-shaped dips; relative stability over a timescale of
30\,days).  We also noted stars that we thought could be CQVs, but
that were more ambiguous with a ``maybe'' flag.

Broadly speaking, the most common false positives for both the Fourier
and dip-counting techniques were eclipsing binaries, spot-induced
variability from rapid rotators, and variability from neighboring,
off-target stars.  Typical false positive rates from our dip-counting
pipeline were 5:1, with \nuniqdipflagged\ unique stars flagged, and
about 20\% being labeled either ``good'' or ``possible'' CQVs; for
the Fourier pipeline, {\bf the rate was X:Y, with ZZZZ unique stars
flagged, and NN\% being labeled ``good'' or ``possible'' CQVs}.

@@

We determined the stellar effective temperature and radii through SED
fitting; we then estimated the masses by interpolating against the
PARSEC v1.2S models \citep{2012MNRAS.427..127B,2014MNRAS.444.2525C}.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f2.pdf}
    \vspace{-0.7cm}
		\caption{
      {\bf CQVs found in the TESS 2-minute data.}
      Phased TESS light curves over one month are shown for \ngoods\
      CQVs.  Gray are raw 2-minute data; black bins to 300 points per
      cycle.  Each panel is labeled by the TIC identifier, the TESS
      sector number, and the period in hours.  Objects are ordered
      such that sources with the most TESS data available are on top
      (see Section~\ref{sec:catalog}).
		}
		\label{fig:cqvs}
	\end{center}
\end{figure*}

@@

As is typical for CQVs \citep{2017AJ....153..152S}, most CQVs in our
catalog did not show infrared excesses in the WISE1-4 bands.
Visually inspecting the SEDs of our entire \ncpvsfound\ star sample,
we labeled two objects as having ``good'' infrared excesses (both W3
and W4), and three as ``possible'' infrared excess.

The two ``good'' IR excesses belonged to ``candidate'' CQVs
TIC~193136669 (TWA~34) and TIC~57830249 (TWA~33).  Both are in the TW
Hydrae association ($\approx$10\,Myr), and have relatively long
periods of 38\,hr and 44\,hr respectively.  In our initial labeling,
we labeled both as ``maybe'' CQVs because the dips in their light
curves did not show the rigid periodicity typical of CQVs;  their
periods were also relatively long.  After labeling them, inspection
of further sectors clarified that both sources are dippers (see plots
in Appendix~\ref{app:vetting}).  In addition, TIC 193136669 has a cold
molecular disk based on observed 1.3\,mm continuum emission and
resolved Keplerian $^{12}{\rm CO}(2-1)$ emission
\citep{2015A&A...582L...5R}.  It was also labeled a dipper by
\citet{2022ApJS..263...14C}; we agree with their designation, but
nonetheless leave it in our catalog as an indication of possible
ambiguities in our search process.  TIC 57830249 (TWA~33) also has
previously detected 1.3\,mm continuum emission
\citep{2015A&A...582L...5R}, suggestive of cold dust grains being
present.  It is also a dipper.
Given these ``errors'', plausible evolutionary connections between CQVs and
dippers are discussed in Section~\ref{subsec:discdippers}.

Our three ``possible'' infrared excesses were TIC~405910546,
289840926, and 244161191.  After a literature search, we concluded
that none have clear evidence for the presence of a disk.
TIC~405910546, in LCC, shows a unique TESS light curve, reminiscent of
a $P$=38\,hr singly-eclipsing binary, except with additional
substructure during each eclipse that resembles the CQV phenomenon
more than any other variability of which we are aware.  TIC~289840926
($\beta$~Pic moving group, $P$=4.8\,hr), show what we believe is a
clear CQV signals, but has no definitive evidence for a large, dusty
disk.  TIC~244161191 (TOI-278), in Columba, also has no
definitive evidence for a large disk.  It is however
``multi-periodic''---in addition to the 7.17\,hr CQV signal, this
source shows a superposed 8.39\,hr sinusoidal signal, probably from an
unresolved neighboring star.

@@

Figure~\ref{fig:evoln} shows ``before'' and ``after'' views of 27 CQVs
for which TESS 120-second cadence observations were available at least
two years apart.  Such a baseline was available for 32 of the
confirmed \ngoods\ CQVs in our catalog; for plotting purposes we show
the brightest 27.  Of these 27 CQVs, a few show clear signs of the
phenomenon in one sector, and marginal or non-existence signs in the
other.  While there is some subjectivity in this assessment, to our
eyes cases for which at least one sector would be flagged as
``ambiguous'' include
TIC~368129164 (Sector 23 might be labeled an EB),
TIC~177309964 (Sector 38 would be simply a rotating star),
TIC~404144841 (Sector 38 looks like a rotating star),
TIC~201898222 (Sector 3 looks like a rotating star),
TIC~144486786 (Sector 32 might be an RS CVn),
and
TIC~38820496 (Sector 28 might be an RS CVn).
TIC~193831684, assessed on a single-sector basis, would probably be
labeled an eclipsing binary---in fact, \citet{2021ApJ...912..123J}
implicitly have already given this source such a label.  However,
based on the the shape evolution between Sectors 13 and 39, it is a
CQV.  One could easily also imagine in cases like TIC~206544316 that
if observed at lower signal to noise, the drastic shape evolution
would not be appreciated.  We emphasize that the periods themselves
were all stable to $<$0.1\%.  Broadly speaking, this shape evolution
suggests that CQVs have an ``on/off'' duty cycle of $\approx$75\%.
This type of correction is likely worth including in population-level
estimates of how intrinsically common CQVs are in the low-mass stellar
population \citep[e.g.][]{2022AJ....163..144G}.

@@

The average period, determined by measuring the PDM peak period over
each sector independently, was $\langle P \rangle = 18.5560$\,hr.  The
range between the maximum and minimum sector-specifics periods was
measured to be about one minute.  Based on this range, the star's
period is stable to at least one minute ($\pm$0.017\,hr) over the
three year baseline.  However, in detail, a period shift of
$\pm$1\,minute would yield major phase drifts over the baseline; that
time interval corresponds to roughly 1/1000$^{\rm th}$ of a period,
and we have observed 1500 cycles.  The achievable period precision,
$\sigma_{\rm P}$, can be estimated as
\begin{equation}
  \sigma_{\rm P} = \frac { \sigma_{\phi} P } { N_{\rm baseline} },
\end{equation}
for $N_{\rm baseline}$ the number of cycles in the observed baseline
and $\sigma_{\phi}$ the phase precision with which any one feature
(e.g.~a dip, or the overall shape of the sinusoidal envelope) can be
tracked.  Assuming $\sigma_{\phi} \approx 0.01$ yields an expected
period precision $\sigma_{\rm P} \approx 0.45\,{\rm sec} \approx
1.2\times 10^{-4}\,{\rm hr}$.  By visually fine-tuning over a grid in
period, we eventually found $P=18.5611 \pm 0.0001$\,hr, which seems to
track certain features in the LP 12-502 light curve over its entire
baseline.

@@

The second dip complex during cycles 248-315 shows the most
substructure.  During e.g.~cycles 283-298, this single complex shows
six local minima.  The deepest dip is sharp: it shows a flux
exercusion of 3.5\% over about 22 minutes (0.02\,$P$), which is the
steepest slope exhibited anywhere in this object's remarkable dataset.
After the sharp dip, there is a roughly exponential fall-off spanning
about a quarter of a period, punctuated by coherent local minima and
maxima which in detail (Appendix~\ref{app:lp}) have slightly longer
periods than the sharp dip.  The sharp leading dip only decreases
during a sudden state-switch at BTJD 2030.7 (cycles 299-315), which
happens immediately during a flare (Figure~\ref{fig:lplc}).  The
trailing dips remain afterwards.

Sectors 53--58 (cycles 1233-1481) are comparatively tame; they showed
only four to six dips per cycle.  Some dips remain stable in depth and
duration over this five month interval.  Other dips grow, like the one
at $\phi = +0.06$ between cycles 1499 and 1481.  Other dips, such as
the one at $\phi = +0.12$ in cycles 1233-1264, disappear entirely.
The most dramatic state switch occurs during cycles 1233-1247, when a
large dip ``switches'' from a depth of 3\% and duration of 3 hours to
a depth of 0.3\%, and a duration of 1 hour.

@@

% 16 px vs 310 px for a full phase
%Based on our SED fitting $R_\star=0.369\,R_\odot$
%star, and from the isochrone mass, $M_\star=0.215\,M_\odot$ 
{\sc Dip durations}---The shortest timescale for any of the {\it
individual} LP 12-502 dips seems to be $\approx$0.06\,$P$ $\approx$
1.08\,hr.  In comparison, using the stellar radius and mass derived in
Section~\ref{subsec:starprops}, the characteristic timescale $T_{\rm
dur} \equiv R_\star P / (\pi a)$ for the transit of a point-source at
corotation is 1.03\,hr ({\bf uncertainties?}).  This means that while
some of the LP 12-502 dips are sufficiently long to require structures
that are extended in orbital azimuth, the durations of other dips are
consistent with effective radii for the occulting material $R_{\rm
eff} \ll R_\star$ .  This implies $a/R_\star \approx 5.8$ for this
material, and so the analogous timescale at the stellar surface is
about six times slower.

@@

\subsection{Dust, or gas?}
We believe that the most likely scenarios to explain the overall CQV
phenomenology are the dust clump scenario, or the prominence scenario
(Figure~\ref{fig:f1}).  Both invoke clumpy material that would be
centrifugally supported at the corotation radius.  The prominence idea
has a longer history, based first on analogy with quiescent
prominences/filaments observed to exist in the solar corona for up to
a few weeks \citep[see e.g.][]{2015ASSL..415.....V}.  In an extrasolar
context, spectroscopic detections of transient Balmer- and
resonance-line absorption seen for stars such as AB~Dor and Speedy~Mic
\citep[e.g.][]{1989MNRAS.238..657C,1993MNRAS.262..369J,2006MNRAS.365..530D,2016MNRAS.463..965L}
led to the interpretation that the data could best be explained by
similar prominence structures: cold, minimally ionized hydrogen clouds
or filaments that scatter chromospheric emission from the star to
produce the observed spectroscopic line variations
\citep[see][]{1989MNRAS.238..657C}.  The (short-term) mechanical
stability of such gas configurations has been demonstrated
\citep{2000MNRAS.316..647F,2022MNRAS.514.5465W}, and the
interpretation of this class of observations seems at least somewhat
secure.

@@

%TODO: improve this to make it less jargony
The difference between the prominence and dust clump scenarios is
essentially only in whether the occulting material of interest is
neutral hydrogen, or dust.  In the phrasing of the ``frozen flux''
condition of ideal rigid field magnetohydrodynamics, the tendency of
both to become trapped at the corotation radius in the equatorial
plane is tied to how of the four relevant forces (gravity, Lorentz,
inertial Coriolis, and inertial centrifugal), the Lorentz and Coriolis
only act perpendicular to field lines, while gravity and the
centrifugal force are in balance at $R_{\rm cr}$
\citep[see][Sec.~2]{2005MNRAS.357..251T}.  The magnetic field strength
is only relevant in this formulation of the system in that we must
have $R_{\rm sonic} < R_{\rm cr} < R_{\rm Alfven}$ in order for closed
loops to exist that can support prominences
\citep{2019MNRAS.482.2853J} {\bf lgb todo: verify \& de-jargonify!}.
In detail however, whether such magnetic fixed points lie in the
equatorial plane, or elsewhere, depends on the star's magnetic field
geometry \citep{2023MNRAS.518.4734S}.

@@

%\subsubsection{Why corotation?}
%All of the prominence models, and the diagrams of where the magnetic
%fixed points exist, enable the fixed points to not be located at
%corotation.
%In addition, the Balmer absorption line measurements have been used to
%argue for material being generally located at positions other than
%just corotation (CHECK Collier-Cameron1989).

%\subsubsection{Does the CQV lifetime inform the mechanism?}
%Independent of the microphysical basis for gas absorption, the
%prominence scenario may present another issue: why don't stars in the
%Hyades and Praesepe show the phenomenon?
%In a broad sense, the prominence scenario argues that an intrinsic
%stellar process is key for generating the CQV phenomenon.
%If this were true, then stars of identical mass, size, and rotation
%periods might be expected to show the phenomenon in equal number,
%simply because the stellar dynamo is generated by fluid
%motions inside the star, which should be identical for stars of the
%same mass, size, and rotation period.
%
%Between ages of $\approx$100\,Myr (AB Dor, Pleiades, Psc-Eri) and
%$\approx$700\,Myr (Praesepe and Hyades), a 0.3\,M$_\odot$ star will
%shrink by $\approx$10\%, from $\approx$0.33\,R$_\odot$ to 
%$\approx$0.29\,R$_\odot$ \citep[e.g.][]{2014AJ....147...82C}.
%This is not a particularly drastic size change.
%Similarly, at 700\,Myr many M-dwarfs have comparably rapid rotation periods 
%to those at 100\,Myr \citep[e.g. Figure~13 of][]{2022AJ....164...80R};
%the mean population does show evidence for spin-down after the
%PMS contraction finishes, but there are still a large number of 
%700\,Myr M-dwarfs with $P_{\rm rot}\lesssim$2\,days, beyond
%which the CQV phenomenon becomes less common.
%
%Broadly speaking -- the absence of old CQVs seems to be
%more plausibly linked to a gradual depletion of dusty detritus from the
%planet formation era than it does to the changing stellar properties
%as the stars gradually finish their pre-main-sequence evolution, and then
%continue to spin rapidly for many hundreds of millions of years.

@@

It is tempting to try to interpret features of the CQV light curves in
terms of the possible presence of close-in exoplanets, or even
planetesimals.  Rocky planets this young would likely have molten
global magma oceans analogous to those that existed on the Earth and
Moon \citep[see][]{2022arXiv220310023L}, and thus would being
undergoing significant outgassing and atmospheric escape.  While a
scenario in which close-in rocky planets serve as a possible source of
dust for the clouds is {\it a priori} plausible, it is currently only
a subject of speculation.

@@

It is amusing that in identifying the two candidate CQV sources in
Section~\ref{subsec:irexcess} with outlying SEDs (TICs~193136669 and
TIC~57830249), we were prompted to reconsider our light curve-based
labeling, to
ultimately conclude that these sources are dippers.  This episode
suggests that could be some overlap between the two classes of
object.  It is also worth emphasizing that our labeling of
e.g.~TIC~57830249 was based on a single sector of TESS data
(Sector~36) when its behavior is relatively periodic and it showed
relative dip depths of a few percent.  However in other TESS sectors
(e.g. Sector~10), this source looks completely different, varying in
apparent flux by a factor of two, with hardly any discernible
periodicity at all.

@@

It is challenging to interpret the significance of such apparent
regularities;  the objects exhibit such a wide range of variability
that for seemingly any ``regularity'' one might posit, it is easy to
come up with counter-example objects which do not follow the trend.
In other words, these half-cycle CQVs could simply be a product of the
human tendency to pattern match; alternatively, they might yield some
important physical significance.

@@

Finally, Figure~\ref{fig:lpriver0} shows ``river plots'' of the
same data, split into similar intervals:
the Sector 18-19 data, 25-26 data, 53 data, and 58-59 data.
State changes are evident in these plots whenever there is a sudden
change in color.
{\bf todo: fix missing data to be different from the flare color}
